---
title: "project1"
author: "jl4756"
date: "9/18/2017"
output:
  pdf_document: default
  html_document: default
---

```{r}
library("rvest")
library("qdap")
library("syuzhet")
library("dplyr")
library("beeswarm")
library("tibble")
library("sentimentr")
library("factoextra")
library("scales")
library("RColorBrewer")
library("RANN")
source("../lib/plotstacked.R")
source("../lib/speechFuncs.R")
```
# read in url
```{r, message=FALSE, warning=FALSE}
### Inauguaral speeches
main.page <- read_html(x = "http://www.presidency.ucsb.edu/inaugurals.php")
inaug=f.speechlinks(main.page)
as.Date(inaug[,1], format="%B %e, %Y")
inaug=inaug[-nrow(inaug),] # remove the last line, irrelevant due to error.

#### Nomination speeches
main.page=read_html("http://www.presidency.ucsb.edu/nomination.php")
nomin <- f.speechlinks(main.page)
nomin<-nomin[-47,] # remove the irrelevant line.

#### Farewell speeches
main.page=read_html("http://www.presidency.ucsb.edu/farewell_addresses.php")
farewell <- f.speechlinks(main.page)
```
# read in list
```{r}
inaug.list=read.csv("inauglist.csv", stringsAsFactors = FALSE)
nomin.list=read.csv("nominlist.csv", stringsAsFactors = FALSE)
farewell.list=read.csv("farewelllist.csv", stringsAsFactors = FALSE)
```
# combine list and url
```{r}
speech.list=rbind(inaug.list, nomin.list, farewell.list)
speech.list$type=c(rep("inaug", nrow(inaug.list)),
                   rep("nomin", nrow(nomin.list)),
                   rep("farewell", nrow(farewell.list)))
speech.url=rbind(inaug, nomin, farewell)
speech.list=cbind(speech.list, speech.url)
```
# write in full text
```{r}
# Loop over each row in speech.list
speech.list$fulltext=NA
for(i in seq(nrow(speech.list))) {
  text <- read_html(speech.list$urls[i]) %>% # load the page
    html_nodes(".displaytext") %>% # isloate the text
    html_text() # get the text
  speech.list$fulltext[i]=text
  # Create the file name
  filename <- paste0("../data/fulltext/", 
                     speech.list$type[i],
                     speech.list$File[i], "-", 
                     speech.list$Term[i], ".txt")
  sink(file = filename) %>% # open file to write 
  cat(text)  # write the file
  sink() # close the file
}
```
# write in Trump's speeches. 
```{r}
speech1=paste(readLines("../data/fulltext/SpeechDonaldTrump-NA.txt", 
                  n=-1, skipNul=TRUE),
              collapse=" ")
speech2=paste(readLines("../data/fulltext/SpeechDonaldTrump-NA2.txt", 
                  n=-1, skipNul=TRUE),
              collapse=" ")
speech3=paste(readLines("../data/fulltext/PressDonaldTrump-NA.txt", 
                  n=-1, skipNul=TRUE),
              collapse=" ")

Trump.speeches=data.frame(
  X...President=rep("Donald J. Trump", 3),
  File=rep("DonaldJTrump", 3),
  Term=rep(0, 3),
  Party=rep("Republican", 3),
  Date=c("August 31, 2016", "September 7, 2016", "January 11, 2017"),
  Words=c(word_count(speech1), word_count(speech2), word_count(speech3)),
  Win=rep("yes", 3),
  type=rep("speeches", 3),
  links=rep(NA, 3),
  urls=rep(NA, 3),
  fulltext=c(speech1, speech2, speech3)
)
speech.list=rbind(speech.list, Trump.speeches)
```

```{r, message=FALSE, warning=FALSE}
sentence.list=NULL
for(i in 1:nrow(speech.list)){
  sentences=sent_detect(speech.list$fulltext[i],
                        endmarks = c("?", ".", "!", "|",";"))
  if(length(sentences)>0){
    emotions=get_nrc_sentiment(sentences)
    word.count=word_count(sentences)
    # colnames(emotions)=paste0("emo.", colnames(emotions))
    # in case the word counts are zeros?
    emotions=diag(1/(word.count+0.01))%*%as.matrix(emotions)
    sentence.list=rbind(sentence.list, 
                        cbind(speech.list[i,-ncol(speech.list)],
                              sentences=as.character(sentences), 
                              word.count,
                              emotions,
                              sent.id=1:length(sentences)
                              )
    )
  }
}
```
# remove non-sentences
```{r}
sentence.list=
  sentence.list%>%
  filter(!is.na(word.count)) 
```
# choose only "Democratic" and "Republican" party to compare.
```{r}
newlist=sentence.list%>%filter(!is.na(Party))
newlist<-rbind(newlist[newlist$Party=="Democratic",],newlist[newlist$Party=="Republican",])
```
# compare only parties
```{r,warning=FALSE}
party.summary<-aggregate(newlist[,13:22],list(newlist$Party),mean)
party.summary=as.data.frame(party.summary)
party.summary$ratio=(party.summary$negative)/(party.summary$positive)
# negative:positive rate, the smaller the better
party.summary$ratio
# the ratio of "Democratic" is smaller than the one of "Republican", which means, the speech of Democratic presidents tend to use positive words.
rownames(party.summary)<-party.summary[,1]
party.summary<-party.summary[,-1]
# create a plot to compare different emotions
{x<-c(1:8)
plot(x,party.summary[1,1:8],type="l",col=1,xaxt = "n")
lines(x,party.summary[2,1:8],type="l",col=2)
legend("topright",c("Democratic","Republican"),lty=1,col=c("black","red"))
axis(1,at=1:8,labels=c("anger","anticipation","disgust","fear","joy","sadness","surprise","trust"))}
```

# compare through president, order by party
```{r}
president.summary<-aggregate(newlist[,13:22],list(newlist$File,newlist$Party),mean)
president.summary=as.data.frame(president.summary)
president.summary$ratio=(president.summary$negative)/(president.summary$positive)
rownames(president.summary)<-president.summary[,1]
president.summary<-president.summary[,-1]
colnames(president.summary)[1]<-c("Party")
# compare ratio in numbers of presidents from different parties.
compare.ratio<-president.summary[order(president.summary$ratio),]
table(compare.ratio$Party[compare.ratio$ratio<0.5])
# compare positive word rates in numbers of presidents from different parties.
compare.pos<-president.summary[order(president.summary$positive),]
table(compare.pos$Party[compare.pos$positive>0.075])
# there isn't too much difference between the number of presidents who likes to use positive words
```

# compare between Terms
```{r}
term.summary<-aggregate(newlist[,13:22],list(newlist$Term),mean)
term.summary=as.data.frame(term.summary)
term.summary$ratio=(term.summary$negative)/(term.summary$positive)
rownames(term.summary)<-term.summary[,1]
term.summary<-term.summary[,-1]
term.summary$ratio # as we can see, presidents tend to use more possitive words when they become a president.
```

# compare between Terms an Parties
```{r}
tp.summary<-aggregate(newlist[,13:22],list(newlist$Term,newlist$Party),mean)
tp.summary=as.data.frame(tp.summary)
tp.summary$ratio=(tp.summary$negative)/(tp.summary$positive)
colnames(tp.summary)[1:2]<-c("Term","Party")

{x<-c(1:4)
plot(x,tp.summary[1:4,13],type="l",col=1,xlab="Term",ylab="ratio",main="Democratic ratio by term",xaxt = "n")
lines(tp.summary[6:7,13],type="l",col=2)
axis(1,at=1:4,labels=c(1,2,3,4))
legend("topright",c("Democratic","Republican"),lty=1,col=c("black","red"))
}
# we can see Democratic presidents tend to use more positive words from their Term1 to Term2, except for president FranklinDRoosevelt, he used more negative words during his 3rd Term because the USA is in WWII, and he used more positive words in his 4th Term because he wants to inspiring citizen.
# However we can see that Republican President tends to use more negative words in their 2nd Term.
```

# compare between type of files.
```{r}
type.summary<-aggregate(newlist[,13:22],list(newlist$type),mean)
type.summary=as.data.frame(type.summary)
type.summary$ratio=(type.summary$negative)/(type.summary$positive)
rownames(type.summary)<-type.summary[,1]
type.summary<-type.summary[,-1]
{x<-c(1:3)
plot(x,tp.summary[1:3,11],type="l",col=1,xlab="Type",ylab="ratio",xaxt = "n")
axis(1,at=1:3,labels=rownames(type.summary)[1:3])
}
# we can see that there is a big decrease of negative words from nomin to inaug, I suppose this may because of pointing out the dark side of America's current situation, and after inaug, in order to make citizen believe that his policy is actually making some progress in improving America's situation, presidents tend to use more possitive words.
typ.summary<-aggregate(newlist[,13:22],list(newlist$type,newlist$Party),mean)
typ.summary=as.data.frame(typ.summary)
typ.summary$ratio=(typ.summary$negative)/(typ.summary$positive)
colnames(typ.summary)[1:2]<-c("Type","Party")
{x<-c(1:3)
plot(x,typ.summary[1:3,13],type="l",col=1,xlab="Type",ylab="ratio",xaxt = "n",ylim=c(min(typ.summary[,13]),max(typ.summary[,13])))
lines(typ.summary[4:6,13],type="l",col=2)
axis(1,at=1:3,labels=c("farewell","inaug","nomin"))
legend("topright",c("Democratic","Republican"),lty=1,col=c("black","red"))
}
```
# choose those that are famous presidents
```{r}
sel.comparison=c("DonaldJTrump","JohnMcCain", "GeorgeBush", "MittRomney", "GeorgeWBush",
                 "RonaldReagan","AlbertGore,Jr", "HillaryClinton","JohnFKerry", 
                 "WilliamJClinton","HarrySTruman", "BarackObama", "LyndonBJohnson",
                 "GeraldRFord", "JimmyCarter", "DwightDEisenhower", "FranklinDRoosevelt",
                 "HerbertHoover","JohnFKennedy","RichardNixon","WoodrowWilson", 
                 "AbrahamLincoln", "TheodoreRoosevelt", "JamesGarfield", 
                 "JohnQuincyAdams", "UlyssesSGrant", "ThomasJefferson",
                 "GeorgeWashington", "WilliamHowardTaft", "AndrewJackson",
                 "WilliamHenryHarrison", "JohnAdams")
```
# First term
```{r, fig.width = 10, fig.height = 10}
par(mar=c(4, 11, 2, 2))
#sel.comparison=levels(sentence.list$FileOrdered)
sentence.list.sel=filter(sentence.list, 
                        type=="nomin", Term==1, File%in%sel.comparison)
sentence.list.sel$File=factor(sentence.list.sel$File)

sentence.list.sel$FileOrdered=reorder(sentence.list.sel$File, 
                                  sentence.list.sel$word.count, 
                                  mean, 
                                  order=T)
beeswarm(word.count~FileOrdered, 
         data=sentence.list.sel,
         horizontal = TRUE, 
         pch=16, col=alpha(brewer.pal(9, "Set1"), 0.6), 
         cex=0.55, cex.axis=0.8, cex.lab=0.8,
         spacing=5/nlevels(sentence.list.sel$FileOrdered),
         las=2, xlab="Number of words in a sentence.", ylab="",
         main="Nomination speeches")
```

# second term
```{r, fig.width = 10, fig.height = 10}
par(mar=c(4, 11, 2, 2))

#sel.comparison=levels(sentence.list$FileOrdered)
sentence.list.sel=filter(sentence.list, 
                        type=="nomin", Term==2, File%in%sel.comparison)
sentence.list.sel$File=factor(sentence.list.sel$File)

sentence.list.sel$FileOrdered=reorder(sentence.list.sel$File, 
                                  sentence.list.sel$word.count, 
                                  mean, 
                                  order=T)

beeswarm(word.count~FileOrdered, 
         data=sentence.list.sel,
         horizontal = TRUE, 
         pch=16, col=alpha(brewer.pal(9, "Set1"), 0.6), 
         cex=0.55, cex.axis=0.8, cex.lab=0.8,
         spacing=1.2/nlevels(sentence.list.sel$FileOrdered),
         las=2, xlab="Number of words in a sentence.", ylab="",
         main="Nomination speeches, 2nd term")
```

#find the longest length of word in each sentence
```{r}
word<-matrix(0,nrow=21326,ncol=max(sentence.list$word.count)+1)
for (i in 1:21326){
  word[i,1:length(nchar(strsplit(as.character(sentence.list$sentences[i]),split="\\, |\\,| |\\:|\\-|\\>")[[1]]))]<-nchar(strsplit(as.character(sentence.list$sentences[i]),split="\\, |\\,| |\\:|\\-|\\>")[[1]])
}
word[,1:124]<-as.numeric(word[,1:124])
rownames(word)<-sentence.list$File

# find the longest word in each sentence.
maxlength.word<-matrix(NA,nrow=nrow(word),ncol=2)
colnames(maxlength.word)<-c("President","max length of word")
maxlength.word[,1]<-sentence.list$File
president<-as.matrix(as.data.frame(table(sentence.list$File))[,1])
maxlength.word[,2]<-apply(word[,1:124],1,max)
table(as.numeric(maxlength.word[,2]))

# find the number of complex words in each president's speeches.
complexword<-matrix(0,ncol=2,nrow=length(president))
complexword[,1]<-as.matrix(president)
for(i in 1:length(president)){
  complexword[i,2]<-sum(as.numeric(word[rownames(word)==president[i],1:124])>=9)
}
complexword.ordered<-complexword[order(as.numeric(complexword[,2])),]
head(complexword.ordered)
tail(complexword.ordered)
# Surprisingly, President Trump is the highest three presidents of using complex words.
```